Epistemic Logics For Knowledge Bases

### 2010-12-15 neighborhood_semantics.tex
factored out completeness theorem

\begin{definition}
  Let $\Phi$ be a set of letters and define the language $\mathcal{L}_N (\Phi)$ as:
  \[ \phi \  : : = \  p \in \Phi \  |
     \  \bot \  | \  \phi \rightarrow \psi
     \  | \  \Box \phi \  | \  A :
     \phi \  | \  K \phi \]
\end{definition}

@@

\begin{proof}
  Assume that $\nvdash \psi$, and define the finitary canonical model
  $\mathbbm{M}= \langle W, V, \mathcal{K}, \mathcal{N}, \mathcal{A} \rangle$
  where:
  \begin{itemizedot}
    \item $W \assign \text{the maximally consistent sets of subformulae of
    $\psi$, closed under single negation}$

    \item $V (p) \assign \{v \in W \  | \  p \in v\}$

    \item $\mathcal{K}_w \assign \{S \in 2^W \  | \  \{v
    \in W \  | \  \phi \in v\} \subseteq S \  \&
    \  K \phi \in w\}$

    \item $\mathcal{N}_w \assign \{S \in 2^W \  | \  \{v
    \in W \  | \  \phi \in v\} \subseteq S \  \&
    \  \Box \phi \in w\}$

    \item $\mathcal{A}_w \assign \{\{\phi\} \  | \  A :
    \phi \in w\}$
  \end{itemizedot}
  Note that in this model, there are neighborhoods for both knowledge and
  belief; more work is necessary to find a model based on this one which
  conforms to the semantics in Definition \ref{neighborhoodmodels}.  For now,
  assume that $K$ is just another modality governed by the neighborhoods in
  $\mathcal{K}_w$.  In both neighborhood functions, our idea is that the
  neighborhoods of $w$ in are supersets $S \supseteq \left\llbracket \phi
  \right\rrbracket$ where $\Box \phi \in w$ or $K \phi \in w$, respectively.

  There are steps of the inductive proof of the Truth Lemma for this structure
  which are not obvious nor obviously documented elsewhere.  Assume
  $\mathbbm{M}, w \vDash \Box \chi$, where the Truth Lemma has been proven to
  hold for $\chi$; we must show $\Box \chi \in w$.  By our assumption there
  is some $\phi$ and $S$ where $\Box \phi \in w$ and $S \supseteq \{v \in W
  \  | \  \phi \in v\}$.  We argue that $\vdash \phi
  \rightarrow \chi$; for if not then by the finitary Lindenbaum lemma there is
  some $u \in W$ where $\phi \in u$ and $\chi \nin u$.  Hence $u \in S$ and
  $\mathbbm{M}, u \nvDash \chi$, a contradiction.  Since $\vdash \phi
  \rightarrow \chi$ then we know that $\vdash \Box \phi \rightarrow \Box \chi$
  by our rules, which means that $\Box \chi \in w$ by maximality.

  By the Lindenbaum Lemma we have that there is some $w \in W$ where
  $\mathbbm{M}, w \nvDash \psi$.  $\mathbbm{M}$ makes true a number of
  properties:
  \begin{enumeratenumeric}
    \item $W$ and $\mathcal{A}_v$ are finite for all $v \in W$, and $\psi \in
    \bigcup \mathcal{A_w}$ only if $\psi$ is a subformula of $\phi$

    \item $\mathcal{K}_w \subseteq \mathcal{N}_w$

    \item If $S \in \mathcal{K}_w$ then $w \in S$

    \item $\mathbbm{M}, v \vDash A : \phi \rightarrow \Box \phi$ for all $v
    \in W$

    \item $\mathcal{N}_w$ and $\mathcal{K}_w$ are closed under supersets
  \end{enumeratenumeric}

  As in the proof of Theorem \ref{completeness1}, we use a bisimulation to
  move to a model where property (3) is strengthened to (3$'$): for all $S \in
  \mathcal{N}_w$, $w \in S$ if and only if $S \in \mathcal{K}'_w$.  We defer
  the reader to {\cite{hansen_bisimulation_2007,pauly_bisimulation_1999}} for
  details regarding bisimulation in neighborhood semantics.  Define
  $\mathbbm{M}' \assign \langle W', V', \mathcal{K}', \mathcal{N}',
  \mathcal{A}' \rangle$ such that:
  \begin{itemizedot}
    \item $W' \assign W \uplus W$ wher.  $l, r$ its associated canonical
    injections and $\theta (v_l) \assign \theta (v_r) \assign v$ is a
    left-inverse of $l, r$

    \item $V' (p) \assign \{v_l, v_r \  | \  v \in V
    (p)\}$

    \item $\mathcal{K}' (v_i) \assign \{S \  | \  v_i \in
    S \  \& \  \theta [S] \in \mathcal{K_v} \}$ where $i =
    l, r$

    \item $\mathcal{N}' (v_i) \assign \mathcal{K}' (v_i) \cup \{S \ 
    | \  v_i \nin S \  \& \  \theta [S] \in
    \mathcal{N_v} \}$ where $i = l, r$

    \item $\mathcal{A}' (v_l) \assign \mathcal{A}' (v_r) \assign \mathcal{A}
    (v)$
  \end{itemizedot}

  If we let $Z \assign \{(v, v_r), (v, v_l) \  | \  v \in
  W\}$, it is straightforward to verify that $Z$ is a neighborhood
  bisimulation.  Hence $\mathbbm{M}, w_l \nvDash \psi$.  Along with (3$'$),
  this model makes true (1), (2), (4).  In fact, in light of ($3'$), we can
  see that this model does not need $\mathcal{K}'$, and conforms to the
  semantics in Definition \ref{neighborhoodmodels}.{\hspace*{\fill}}

  As in previous proofs, let $\Lambda$ be the proposition letters occuring in
  $\psi$ and define an injection $\iota : W' \hookrightarrow \Phi \backslash
  \Lambda$, assigning a nominal to every world in $\mathbbm{M}'$.  Define
  $\mathbbm{M}'' \assign \langle W'', V'', \mathcal{N}'', \mathcal{A}''
  \rangle$ such that:

  \begin{center}
    \begin{tabular}{lllll}
      $\bullet$ & $W'' \assign W'$ &  & $\bullet$ & $V'' (p) \assign \left\{
      \begin{array}{ll}
        V' (p) & p \in \Lambda\\
        \{v\} & p = i (v)\\
        \varnothing & o / w
      \end{array} \right.$\\
      $\bullet$ & $\mathcal{N}''_w \assign \mathcal{N}'_w$ & {\hspace{3em}} &
      $\bullet$ & $\mathcal{A}''_v \assign \{\partial_v S \  |
      \  S \in \mathcal{N}'_w \}$
    \end{tabular}
  \end{center}

  Where $\partial_v S \assign \{\phi \  | \  \phi \in
  \bigcup \mathcal{A}_v', \text{$\phi$ is a subformula of $\psi$ and}
  \  S \subseteq \left\llbracket \phi
  \right\rrbracket^{\mathbbm{M}'} \} \cup \{\neg \iota (u) \  |
  \  u \nin S\}$.

  As before, this model agrees with $\mathbbm{M}'$ on all subformulae of
  $\psi$, using the semantics in Definition \ref{neighborhoodmodels}, hence
  $\mathbbm{M}'', w_l \nvDash \phi$. \

  All that is left is to show \tmtextbf{NCSQ} and \tmtextbf{NSND} from three
  new properties:
  \begin{enumerateroman}
    \item For all words $v$ and all $X \in \mathcal{A}_v''$, $X$ is finite and
  .  $X = \partial_v S$ for some $S \in \mathcal{N}_v''$

    \item For all $v$ and all $S \in \mathcal{N}_v$, $u \in S$ if and only if
    $\mathbbm{M}'', u \vDash \bigwedge \partial_v S$

    \item The logic presented in Table \ref{logic3} is sound for
    $\mathbbm{M}''$
  \end{enumerateroman}
  Along with the deduction theorem for $\vdash$, these properties establish
  \tmtextbf{NCSQ} for this model using the logic of Table \ref{logic3} itself:

  \begin{align*}
    \text{$\exists X \in \mathcal{A''_v} \text{ s.t. } \tmop{Th}
    (\mathbbm{M}'') \cup X \vdash \phi$} & {\Longleftrightarrow}\text{$\exists
    S \in \mathcal{N''_v} \text{ s.t. } \tmop{Th} (\mathbbm{M}'') \cup
    \partial_v S \vdash \phi$}\\
    & {\Longleftrightarrow}\text{$\exists S \in \mathcal{N''_v} \text{ s.t. }
    \tmop{Th} (\mathbbm{M}'') \vdash \bigwedge \partial_v S \rightarrow
    \phi$}\\
    & {\Longleftrightarrow}\text{$\exists S \in \mathcal{N''_v} \text{ s.t. }
    \bigwedge \partial_v S \rightarrow \phi \in \tmop{Th} (\mathbbm{M}'')$}\\
    & {\Longleftrightarrow}\text{$\exists S \in \mathcal{N''_v} \text{ s.t. }
    \left\llbracket \bigwedge \partial_v S \right\rrbracket \subseteq
    \left\llbracket \phi \right\rrbracket$}\\
    & {\Longleftrightarrow}\text{$\exists S \in \mathcal{N''_v} \text{ s.t.
    $\forall u \in W$ if } \mathbbm{M}, u \vDash \bigwedge \partial_v S \text{
    then } \mathbbm{M}, u \vDash \phi$}\\
    & {\Longleftrightarrow}\text{$\exists S \in \mathcal{N''_v} \text{ s.t. }
    \mathbbm{M}, u \vDash \phi \text{ for all } u \in S$}\\
    & {\Longleftrightarrow}{\mathbbm{M}},v{\vDash}\Box{\phi}
  \end{align*}

  \tmtextbf{NSND} follows from \tmtextbf{NCSQ} and the semantics of Definition
  \ref{neighborhoodmodels}.
\end{proof}

>>>>

\begin{definition}
  Let $\Phi$ be a set of letters and define the language $\mathcal{L}_N (\Phi)$ as:
  \[ \phi\ : : =\ p \in \Phi\ |
    \ \bot\ |\ \phi \rightarrow \psi
    \ |\ \Box \phi\ |\ A :
     \phi\ |\ K \phi \]
\end{definition}

@@

\input{completeness5.tex}
